Live Functional Programming with Typed Holes

### 2018-03-22 examples.tex

\HazelnutLive{} programs are written in OCaml \cite{leroy03:_ocaml}, using the
alternative Reason syntax \cite{reason-what} extended to support holes.
%
As shown in \autoref{fig:grades-example}, a \HazelnutLive{} notebook is a
literate program \cite{knuth1984literate}: rich text comments appear at the top
level and code appears within cells embedded into these comments.
%
The reader is encouraged to work through these examples in our prototype
implementation or watch video demonstrations, both available in
\suppMaterials{}.\footnote{
%
The programs in this section are written using minor syntactic conveniences that
are not currently supported in our implementation.
%
For example, the implementation of recursive types in our current implementation
requires explicit \texttt{fold} and \texttt{unfold} expressions; these can be
inferred in standard ways (by pairing them with constructors and deconstructors
for algebraic datatypes) in future work.
%
More details about other minor syntactic differences are described in
\suppMaterials{}.
%
}

>>>>

The screenshot at the top of \autoref{fig:grades-example} shows a mockup of
\Hazel{}, a work-in-progress programming environment that we are building on top
of our dynamic semantics.
%
\Hazel{} programs are written in OCaml \cite{leroy03:_ocaml}, using the
alternative Reason syntax \cite{reason-what} extended to support holes.
%
On the left side of the window is the \emph{edit action panel}, which
provides a user interface for performing structured edits to the program
displayed in the center of the window.
%
On the right side of the window is the \emph{environment panel}, which provides
static and dynamic information about the expression in the program currently in
focus (\ie{}~under the cursor).

Our current implementation, a reference implementation of the \HazelnutLive{}
calculus, provides the edit action, code, and environment panels as shown in
\autoref{fig:grades-example}, but does not yet provide support for many of the
programming constructs and syntactic sugar described in the examples in this
section.
%
\autoref{sec:implementation} describes the status of \HazelnutLive{} in more
detail.
%
%
%% As shown in \autoref{fig:grades-example}, a \HazelnutLive{} notebook is a
%% literate program \cite{knuth1984literate}: rich text comments appear at the top
%% level and code appears within cells embedded into these comments.
%
The reader is encouraged \rkc{(maybe not)}
to work through the following examples in our prototype
implementation or watch video demonstrations, both available in
\suppMaterials{}.

### 2018-03-22 misc.tex

>>>>

\subsection{Raw Material for Related Work}

\begin{itemize}
	\item Hazelnut paper
	\item simply typed underdeterminism paper
	\item work on partial evaluation and staging (incl. connections to modal logic -- ``modal analysis of staged computation'')
	\item symbolic evaluation
	\item full beta reduction
	\item Agda
	\item Idris
	\item GHC holes
	\item CMTT
	\item gradual typing (and dynamic typing)
		\begin{itemize}
			\item siek and taha paper
			\item snapl15 paper
			\item gradualizer paper
			\item maybe other things, e.g. several papers by ron garcia
		\end{itemize}
	\item DuctileJ stuff -- \url{https://homes.cs.washington.edu/~mernst/pubs/ductile-icse2011.pdf}
	\item OLEG from McBride's thesis
	\item Visual Studio (and others) support for edit-and-resume
	\item Scratch lets you just skip over statement holes
	\item prior work on confluence for the lambda calculus
	\item work on debuggers that allow you to inspect environments
		\begin{itemize}
		\item might be something relevant in the paper ``A Debugger for Standard ML'' 
		\item "Visualizing the evaluation of functional programs for debugging" by Whitington and Ridge
		\item "A lightweight interactive debugger for Haskell'' and ``Multiple-View Tracing for Haskell: a New Hat'' might be relevant
		\item ocamli -- \url{https://github.com/johnwhitington/ocamli}
		\item Better supporting debugging aids learning a novel programming language. -- Scaffidi at VLHCC 2017
		\item quote from Wadler in ``Why no one uses functional languages'':
			\begin{quote}
			“...there are few debuggers or
profilers for strict [functional] languages, perhaps because constructing them is not considered
research. This is a shame, since such tools are sorely needed, and there remains much of
interest to learn about their construction and use.
\end{quote}
		\end{itemize}
	\item papers that show up in a search for ``typed holes'' -- \url{https://scholar.google.com/scholar?hl=en&as_sdt=0%2C39&q=%22typed+holes%22&btnG=}
	\item maybe also search for ``partial programs'' 
	\item roly pererra work on program slicing
	\item should say something about how holes show up in program synthesis under various names (what?) 
	\item ``Achieving flexibility in direct-manipulation programming environments by relaxing the edit-time grammar'' -- \url{http://ieeexplore.ieee.org/document/1509511/}
	\item ``Call-by-value is dual to call-by-name'' might be relevant? \url{http://homepages.inf.ed.ac.uk/wadler/papers/dual/dual.pdf}
	\item mention how unspecified evaluation order is something that people do when talking about parallelism?
\end{itemize}
